Efficiently summarising relationships in large samples: a general duality between statistics of genealogies and genomes

### 2019-09-06 treestats_paper.tex
does the bit on variance need further exposition?
comment re: results seemingly w/o a figure?

@@

As another example, we simulated an ``admixture'' scenario:
a first population of size $N=1000$ splits into two of equal size,
then after $N$ generations, the second population splits,
after another $N$ generations, the third population splits again,
and for the final $N$ generations populations 2 and 3
have per-capita migration rates of $4/N$ to each other.
In this situation, a significantly negative $f_4(1,2;3,4)$ is expected,
which we indeed find, with a mean value of around -700.
Using various sample sizes, mutation rates, and window sizes, we then
calculated this $f_4$ statistic in windows along a 100Mb genome,
and show the standard deviation of both ``site'' and ``branch'' statistics across windows
in Figure \ref{fig:introf4}.
Since the genome is uniform (no selection or variation in recombination or mutation),
this standard deviation is a measure of noisiness.
As expected, site statistics are noisier than branch statistics,
by a factor that depends mostly on the ratio of mutation to recombination rates.
The results suggest that branch statistics would provide substantially better resolution
at small scales along the genome, especially if mutation rate is lower than recombination rate.
However, in practice imperfect estimation of the tree sequence
would introduce additional noise,
so it remains to be seen if the improvement could be made in practice.

>>>>

\krt{will this section go further?  It just kind of ends...}

@@

As another example, we simulated an ``admixture'' scenario:
a first population of size $N=1000$ splits into two of equal size,
then after $N$ generations, the second population splits,
after another $N$ generations, the third population splits again,
and for the final $N$ generations populations 2 and 3
have per-capita migration rates of $4/N$ to each other.
\krt{It is not clear what figure this bit of text refers to?  Is this out of date?}
In this situation, a significantly negative $f_4(1,2;3,4)$ is expected,
which we indeed find, with a mean value of around -700.
Using various sample sizes, mutation rates, and window sizes, we then
calculated this $f_4$ statistic in windows along a 100Mb genome,
and show the standard deviation of both ``site'' and ``branch'' statistics across windows
in Figure \ref{fig:introf4}.
Since the genome is uniform (no selection or variation in recombination or mutation),
this standard deviation is a measure of noisiness.
As expected, site statistics are noisier than branch statistics,
by a factor that depends mostly on the ratio of mutation to recombination rates.
The results suggest that branch statistics would provide substantially better resolution
at small scales along the genome, especially if mutation rate is lower than recombination rate.
However, in practice imperfect estimation of the tree sequence
would introduce additional noise,
so it remains to be seen if the improvement could be made in practice.
